---
##title: "Reproducible Research: Peer Assessment 1"
## * author: "Jesse Sharp"
##date: "August 15, 2015"
##output: html_document
##keep_md: true
---

## Clone the repo
## https://github.com/SharpStatSci/RepData_PeerAssessment1.git

### Loading and preprocessing the data

The first step in the analysis is to load the data into R and take a look at it.
After examining the basic properties of the file we can move on to cleaning or
modifying the dataset if necessary.

```{r}
# set the correct working directory
# we assume data is already downloaded into a data folder

getwd()
setwd("C:/StatWare/Rprog/RepData/project1")

actdata <- read.csv("C:/StatWare/Rprog/RepData/project1/data/activity.csv")

# check the basic properties of the data

head(actdata)
tail(actdata)
str(actdata)
summary(actdata)

sum(is.na(actdata$steps))
sum(!is.na(actdata$steps))

sum(is.na(actdata$date))
sum(!is.na(actdata$date))

sum(is.na(actdata$interval))
sum(!is.na(actdata$interval))
```

### What is mean total number of steps taken per day?

For this section we are asked to plot a histogram of the mean number of steps
per day. In addition, compute the mean and median steps per day. 

```{r}
actdata2 <- actdata[!is.na(actdata$steps),]
summary(actdata2)

actday <- aggregate(actdata2$steps, by=list(actdata2$date), mean)

# mean and median after removing NA
mean(actday$x)
median(actday$x)
```


```{r}
plot1 <- function() {
    par(mfrow = c(1, 2))
    hist(actdata2$steps)
    hist(actday$x)
}
plot1
```

### What is the average daily activity pattern?
```{r}
# steps by interval, interval with max steps

actint <- aggregate(actdata2$steps, by=list(actdata2$interval), mean)
head(actint)

plot(actint$Group.1, actint$x, type="l", xlab = "Interval", ylab = "Average Steps", col="blue", main="Ave Steps by Interval")

summary(actint)
actint[actint$x>=100,] #interval 835
```
### Imputing missing values

```{r}
```

### Are there differences in activity patterns between weekdays and weekends?

```{r}
```
